---
title: "Plots for paper"
author: "Katarina Mandić"
date: "`r format(Sys.time(), '%d %b, %Y')`"
output:
  html_document:
    css: style.css
---
 

# Description
Plotting for the paper

# Libraries
```{r Libraries, message=FALSE}
library(ggplot2)
library(DESeq2)
library("pheatmap")
library("RColorBrewer")
library(dplyr)
library(gprofiler2)
library(rprojroot)

# Leads to the project directory
project_dir <- find_rstudio_root_file()

source(file = file.path(project_dir,"src", "utils.R"))

genes_dir <- file.path(project_dir, "data","input", "genes.hg19.ensembl.rds")
```

# Load data
```{r}
mRNA_dds <- readRDS(file = file.path(project_dir, "data/processed/mRNA_dds.rds"))
                      
mir_dds <- readRDS(file = file.path(project_dir, "data/processed/mir_dds.rds"))
rnb_set <- readRDS(file = file.path(project_dir, "data/processed/rnb_set.rds"))

# This table is the one for the paper, but contains all samples so its usefull
samples_ppr <- readxl::read_xlsx(file.path(project_dir, "data/input/samples.xlsx"))

# Full annotation
annot_full <- readxl::read_xlsx(file.path(project_dir, "data/raw/hnscc_annot.xlsx"))
```


# PCA plot on mRNA data
```{r}
mRNA_rld <- DESeq2::rlog(mRNA_dds)
```


```{r}
# selecting data for plotting pca
mRNA_pcadata <- 
  DESeq2::plotPCA(mRNA_rld, 
                  intgroup = c("Hist", "HPV", "Tissue"),
                  returnData = TRUE) %>%
  dplyr::rename(Sample ="Hist") %>%
  mutate(HPV = ifelse(HPV==1, "P","N"))
# selecting variance
percentVar <- round(100 * attr(mRNA_pcadata, "percentVar"))


# Color for PCA 
col_pca <- c(
  # Dark red
  brewer.pal(8, "Reds")[7],
  brewer.pal(8, "Blues")[7]
)


# plotting pca data
ggplot(mRNA_pcadata, aes(x = PC1, y = PC2, color = Tissue, shape = HPV, fill = Sample)) +
  geom_point(size =4,
             stroke =1.2) +
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
   theme(#panel.grid = element_blank(), 
         # panel.background = element_rect(fill = brewer.pal(3,"Greys")[1]),
     panel.background = element_rect(fill = "white"),
         panel.border = element_rect(fill = "transparent"),
         panel.grid = element_line(color = brewer.pal(5,"Greys")[1]),
         legend.key=element_blank(),
         axis.title=element_text(size=18),
         legend.text=element_text(size=18),
         legend.title=element_text(size=18))+
  coord_fixed() +
  #ggtitle("PCA with rlog data")+
  scale_shape_manual(values = c(21, 22) )+
  scale_color_manual(values = c("black", brewer.pal(5,"Greys")[3]) )+
  scale_fill_manual(values= col_pca)+
  guides(fill=guide_legend(override.aes=list(shape=21)))+
  guides(color=guide_legend(override.aes=list(shape=21))) +
  guides(fill=guide_legend(override.aes=list(shape=21,fill=c(col_pca))))

  #geom_text_repel(aes(label = colData(mRNA_rld)$Sample.ID), size =3)
  #geom_text_repel(aes(label = colData(mRNA_rld)$GRCE.NAME), size =3)

```

Saving to file
```{r, eval=FALSE}
ggsave(filename = file.path(project_dir, "data/output/figures/PCA_mRNA.png"))
```

# miR distance
Since I made a presentable PCA plot for mRNA data I will first make sample distance plot for miR
```{r }
# Variance stabilizng 
mir_rld <- rlog(mir_dds)
# calculating sample distance matrix
sampleDist_mir <- stats::dist(t(assay(mir_rld)), method = "euclidean")
sampleDist_mir <- as.data.frame(as.matrix(sampleDist_mir))
```



```{r}
colors <- colorRampPalette( rev(brewer.pal(9, "Oranges")) )(255)

annot_col <- list(
  # Tumor colors
  Hist = setNames(c(brewer.pal(8, "Reds")[7],  brewer.pal(8, "Blues")[7]),
                   c("cancer", "control")),
  # HPV colors
  HPV = setNames( c("forestgreen","gold"), 
                  c("P","N")),
  # Tissue colors
  Tissue = setNames(c(brewer.pal(8, "Dark2")[3], brewer.pal(8, "Dark2")[7]),
                    c("OP","O")),
    # Smoking drinking colors
  Smoking_Drinking = setNames( c("forestgreen","gold", "white"), 
                  c("0","1", "?")),
  Age = brewer.pal(9, "Greys"),
  Gradus = setNames(brewer.pal(9, "Reds")[c(1,3,5,7,9)],
                    c("nema nalaza", "0", "1", "2", "3"))
)

annot_data <- colData(mir_rld) %>% 
  as.data.frame() %>%
  dplyr::select(Hist, Tissue, HPV) %>%
  # changing normal to control in Hist
  mutate(Hist = ifelse(Hist == "normal", "control", "cancer"))

# additional annotations
annot_data <-
  left_join(tibble::rownames_to_column(annot_data), 
            # first group of annot
          dplyr::select(annot_full, Sample_ID, Smoking_Drinking, Age, Gradus),
          # second group
          # select(annot_full, Sample_ID, Stage, Smoking_Drinking, Gradus),
          by = c("rowname" = "Sample_ID")) %>%
  tibble::column_to_rownames("rowname")


pheatmap(sampleDist_mir,
         # clustering_distance_rows = sampleDist_mir,
         # clustering_distance_cols = sampleDist_mir,
         cluster_rows = TRUE,
         cluster_cols = TRUE,
         col = colors,
         treeheight_col = 0,
         treeheight_row = 0,
         #annotation_row = annot_data,
         annotation_col = annot_data,
         annotation_colors = annot_col,
         show_colnames = TRUE,
         show_rownames = FALSE,
         annotation_names_col = FALSE)
```

Saving to file
```{r, eval=FALSE}
pheatmap(sampleDist_mir,
         # clustering_distance_rows = sampleDist_mir,
         # clustering_distance_cols = sampleDist_mir,
         cluster_rows = TRUE,
         cluster_cols = TRUE,
         col = colors,
         treeheight_col = 0,
         treeheight_row = 0,
         #annotation_row = annot_data,
         annotation_col = annot_data,
         annotation_colors = annot_col,
         show_colnames = TRUE,
         show_rownames = FALSE,
         annotation_names_col = FALSE,
         #fontsize = 20,
         #width = 10,
         #cellwidth = 25,
         filename = file.path(project_dir, "data/output/figures/extra_annot_sampleDistance_miR.png"))
```


# DNAm distance
```{r}
# Novariance stabilizarion?
# calculating sample distance matrix
sampleDist_dnam <- stats::dist(t(meth(rnb_set)))
sampleDist_dnam <- as.data.frame(as.matrix(sampleDist_dnam))
```

```{r}
colors <- colorRampPalette( rev(brewer.pal(9, "Greens")) )(255)
# I have a full list in mir
# annot_col <- list(
#   # Tumor colors
#   Hist = setNames(c(brewer.pal(8, "Reds")[7],  brewer.pal(8, "Blues")[7]),
#                    c("cancer", "control")),
#   # HPV colors
#   HPV = setNames( c("forestgreen","gold"), 
#                   c("P","N")),
#   # Tissue colors
#   Tissue = setNames(c(brewer.pal(8, "Dark2")[3], brewer.pal(8, "Dark2")[7]),
#                     c("OP","O")),
#   # Smoking drinking colors
#   Smoking_Drinking = setNames( c("forestgreen","gold", "pink"), 
#                   c("0","1", "?"))
# )



annot_data <- pheno(rnb_set) %>% 
  as.data.frame() %>%
  dplyr::mutate(Hist = ifelse(Sample_group == "Cancer", "cancer", "control")) %>%
  dplyr::select(Sample_ID, Hist, HPV) %>%
  # Adding tissue locations from sample_ppr table 
  left_join(dplyr::select(samples_ppr,ID, Tissue = "Location") %>%
              mutate(ID=tolower(ID)),
            by = c("Sample_ID" = "ID")) %>%
  dplyr::select(Hist, Tissue, HPV)

rownames(annot_data) <- pheno(rnb_set)$Sample_ID



# additional annotations
annot_data <-
  left_join(tibble::rownames_to_column(annot_data), 
            # first group of annot
          dplyr::select(annot_full, Sample_ID, Smoking_Drinking, Age, Gradus) %>%
          # second group
          # select(annot_full, Sample_ID, Stage, Smoking_Drinking, Gradus) %>%
            mutate(Sample_ID = tolower(Sample_ID)),
          by = c("rowname" = "Sample_ID")) %>%
  tibble::column_to_rownames("rowname")

pheatmap(sampleDist_dnam,
         # clustering_distance_rows = sampleDist_dnam,
         # clustering_distance_cols = sampleDist_dnam,
         col = colors,
         treeheight_col = 0,
         treeheight_row = 0,
         #annotation_row = annot_data,
         annotation_col = annot_data,
         annotation_colors = annot_col,
         show_colnames = TRUE,
         show_rownames = FALSE,
         annotation_names_col = FALSE)

```

Saving to file
```{r, eval=FALSE}
pheatmap(sampleDist_dnam,
         # clustering_distance_rows = sampleDist_dnam,
         # clustering_distance_cols = sampleDist_dnam,
         col = colors,
         treeheight_col = 0,
         treeheight_row = 0,
         #annotation_row = annot_data,
         annotation_col = annot_data,
         annotation_colors = annot_col,
         show_colnames = TRUE,
         show_rownames = FALSE,
         annotation_names_col = FALSE,
         #fontsize = 20,
         #width = 10,
         #cellwidth = 25,
         filename =file.path(project_dir, "data/output/figures/extra_annot_sampleDistance_dnam.png"))
```


# DE expression gsea

## Loading data
```{r}
mRNA_res <- readRDS(file.path(project_dir, "data/processed/mRNA_res.rds"))

```


```{r}
mRNA_gost <- as.data.frame(mRNA_res)%>%
  filter(padj < 0.05) %>%
  rownames() %>%
  unique()


# Using only part of sources
GSEA_SOURCES <- c("GO:BP","KEGG","REAC")


# Call on the gost 
mRNA_gost <- 
  gost(mRNA_gost,
       organism= "hsapiens",
       exclude_iea=TRUE,
       domain_scope="annotated",
       ordered_query=F,
       sources=GSEA_SOURCES)

# Change query name
mRNA_gost$result$query <- ""

highlighted_gsea_terms <- read.csv(file.path(project_dir,"data/input/GSEA_terms.txt"),
                                   header = F)[,1]

#Plot and table of significant results
gostplot(mRNA_gost ,capped = F, interactive = T)

publish_gostplot(gostplot(mRNA_gost ,capped = F, interactive = F),
                 highlight_terms = highlighted_gsea_terms)

```

```{r, eval=FALSE}
publish_gostplot(gostplot(mRNA_gost ,capped = F, interactive = F),
                 highlight_terms = highlighted_gsea_terms,
                 filename = file.path(project_dir, "data/output/figures/gost_plot_mRNA_highlighted.png"))
```


# mir targets histograms
will not be doing that
## Loading data
```{r, eval=FALSE}
targ_mir_mRNA <- 
  readRDS( "../02-Regulation_targets/outputs/RDS/targ_mir_mRNA.rds")

cor_mir <-
  readRDS("../02-Regulation_targets/outputs/RDS/cor_mir.rds")
```

## plot
```{r, eval=FALSE}
mir_freq <- 
  table(cor_mir$mir) %>%
  as.data.frame()


ggplot(mir_freq, aes(Freq))+
  geom_histogram()+
  stat_bin(binwidth = 1, bins = 50)
```
# Mir mRNA scatterplots

```{r}

```


# Session Info
```{r Session info}
sessionInfo()
```
